---
title: "02 - Why and How to Invert Edges"
author: "Otho Mantegazza"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{02 - Why and How to Invert Edges}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

# Set up

```{r, message = FALSE}
library(igraph)
library(ggraph)
library(magrittr)
library(ptrapr)
library(tidygraph)
```

# Issue - direction of the edges in ptrapr output

Sometimes, the direction of the edges in the graphs produced with ptrap is inverted.

Every edge correspond to the part of a branch that connect one node to another. By convention each edge should be directed outward: pointing from the node closer to the base of the panicle to the one further away.

You can understand this better by looking at the panicle graphs: we provide two panicle graph objects:

- One named `good_panicle` in which every edge points outward.
- One named `inverted_panicle` in which the edges on the main rachis axis point inward.

We can see this comparing the graphs. 

The `good_panicle` graph has no inverted edges:
 
```{r}
good_panicle %>% plot_panicle()
```

Compare the previous graph with the one below, representing the `inverted_panicle` object. In the graph below, the edges on the main axis are inverted.

```{r}
inverted_panicle %>% plot_panicle()
```

# Invert edges with invert_edges()

If you find a panicle with inverted edges on the main axis, you can use the function `invert_edges()` to revert them.

```{r}
tst <- inverted_panicle %>% invert_edges()
tst %>% plot_panicle()


inverted_panicle %>%
  invert_edges(check_before = TRUE) %>%
  plot_panicle()

good_panicle %>%
  invert_edges(check_before = TRUE) %>%
  plot_panicle()
``` 


# Details

- The main axis is defined as all the edges between nodes with both type "Generating" or "Primary"


