---
title: "05 - Extract and Save Basic Panicle Features"
author: "Otho Mantegazza"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{05 - Extract and Save Basic Panicle Features}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

# Setup

```{r, message=FALSE}
library(ptrapr)
library(magrittr)
library(igraph)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
```

# Extract features

You can easily extract features from a panicle object, such as number of branches and spikelets. Any panicle object is build upon an `igraph` object, so you can use functions from `igraph` to extract features from panicles.  Or learn a few tricks below.

We can start from the object `good_panicle`.

```{r}
good_panicle %>% plot_panicle()
```

# Features from one panicle with vertex_attr()

The function `vertex_attr()` from igraph extracts the attributes from all the nodes (vertexes) of a panicle.

```{r}
nodes_df <- 
  good_panicle %>% 
  vertex_attr() %>% 
  # The output of vertex_attr() is a list,
  # but  we can manipulate it more easily 
  # if we turn it into a data frame
  as_tibble()
```

This is the output of `vertex_attr()`, the column `type` stores the type of each node (primary, secondary, etc.):

```{r}
# use head() to take only the first 6 rows
nodes_df %>% head() %>% knitr::kable()
```

## These are the panicle features

We can count the feature with dplyr's `count()`. These are the features from one panicle.

```{r}
pan_feats <- 
  nodes_df %>% 
  group_by(type) %>% 
  count()
```

```{r, echo = FALSE}
pan_feats %>% knitr::kable()
```

# loop over many panicles with map() and reduce()

## load many panicles

As in [the previous vignette](04-visualize-many-panicles-loop-r.html), you can parse many panicles and put them in a list. Check that vignette to for a torough explaination of these steps.

```{r, echo = FALSE}
path_to_folder <- system.file(package = "ptrapr") %>%
  paste0("/extdata") 


# First record the names of the files
# substitute path_to_folder with the path to your folder
panicle_paths <- 
  list.files(path_to_folder) %>%
  # remove the extension (variable part)
  # each panicle has two files with different extension
  str_remove(".ricegr|.ricepr") %>%
  # store the name of the files without the extension
  unique()

panicle_paths
# Store file names in named object, to use them later
panicle_paths <-
  panicle_paths %>% 
  {set_names(x = paste(path_to_folder, ., sep = "/"), nm = .)}

# Parse them in a list, using file name as list names
panicle_list <- 
  panicle_paths %>%
  map(~read_full_panicle(.)) 


# Then correct the edges on the main axis with invert_edges()
panicle_list <- 
  panicle_list %>% 
  map(~invert_edges(., check_before = TRUE))


names(panicle_list)
```

Afterwards you can `map()`, `bind_rows()` and `spread()` to collect all the features in a database.

```{r, eval = FALSE}
panicles_feature <- 
  panicle_list %>% 
  map(~vertex_attr(.) %>%
        as_tibble() %>% 
        group_by(type) %>%
        count()) %>%  #%>% 
        # {set_names(x = .$n, nm = .$type)}) %>% 
  bind_rows(.id = "id") %>% 
  spread(key = "type", value = "n")

panicles_feature
```

The last step `spread()` is not necessary, depends if need like your data in [long or wide format](https://r4ds.had.co.nz/tidy-data.html#spreading-and-gathering).

Let me know if you want me to code these steps in a function.

# More about  igraph and tidygraph

Check the [igraph website](https://igraph.org/r/) for a details about it, also [tidygraph](https://www.data-imaginist.com/2017/introducing-tidygraph/) can be useful.
