---
title: "05 - Extract and Save Basic Panicle Features"
author: "Otho Mantegazza"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{05 - Extract and Save Basic Panicle Features}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

# Setup

```{r, message=FALSE}
library(ptrapr)
library(magrittr)
library(igraph)
library(dplyr)
library(purrr)
```

# Extract features

You can easily extract features from a panicle object, such as number of branches and spikelets. Any panicle object is build upon an `igraph` object, so you can use functions from `igraph` to extract features from panicles.  Or learn a few tricks below.

We can start from the object `good_panicle`.

```{r}
good_panicle %>% plot_panicle()
```

# Features from one panicle with vertex_attr()

The function `vertex_attr()` from igraph extracts the attributes from all the nodes (vertexes) of a panicle.

```{r}
nodes_df <- 
  good_panicle %>% 
  vertex_attr() %>% 
  # The output of vertex_attr() is a list,
  # but  we can manipulate it more easily 
  # if we turn it into a data frame
  as_tibble()
```

This is the output of `vertex_attr()`, the column `type` stores the type of each node (primary, secondary, etc.):

```{r}
# use head() to take only the first 6 rows
nodes_df %>% head() %>% knitr::kable()
```

## These are the panicle features

We can count the feature with dplyr's `count()`. These are the features from one panicle.

```{r}
pan_feats <- 
  nodes_df %>% 
  group_by(type) %>% 
  count()
```

```{r, echo = FALSE}
pan_feats %>% knitr::kable()
```

# loop over many panicles with map() and reduce()

```{r}
# put all the panicles into a list
panicle_list <- list(good_panicle,
                     inverted_panicle %>% invert_edges())
```

```{r}
# loop with map
panicle_features <- 
  panicle_list %>% 
  map(~vertex_attr(.) %>%
        as_tibble() %>%
        group_by(type) %>% 
        count()) %>% 
  reduce(left_join, by = "type")
```

# More about  igraph and tidygraph

Check the [igraph website](https://igraph.org/r/) for a details about it, also [tidygraph](https://www.data-imaginist.com/2017/introducing-tidygraph/) can be useful.
