---
title: "01 - Load Panicles into R"
author: "Otho Mantegazza"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01 - Load Panicles into R}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r, message=FALSE}
library(ptrapr)
library(ggraph)
library(magrittr)
library(igraph)
# library(tibble)
library(dplyr)
# library(tidygraph)

# ggplot2 theme minimal fits better with graphs
theme_set(theme_minimal())
```



# Read Panicle XML files

The software [ptrap](https://bioinfo.ird.fr/index.php/resources/49-uncategorised/102-p-trap) scans pictures of rice panicles and outputs their architecture in two XML files:

A structure file:
: One file, with extension `.ricepr`, stores the xy coordinates of each
  element of a panicle besides spikelets/grains.
  This file also stores the topology of those panicle elements in a
  directed graph.
  
A grain file:
: A second file, with extension `.ricegr`, store the xy coordinates of each spikelet.

We have included examples of such files in this package. You can access these examples **online* [here](https://github.com/othomantegazza/ptrapr) or locally in the `ptrapr/extdata` folder [^find.package]

[^find.package]: Locate your local package folder with the function:: `find.package("ptrapr")`
 
## Read Panicle Structure with `read_panicle()`

The structure and topology of the panicle is stored in the `.ricepr` XML file.

You can test the `read_panicle()` function on one of the `.ricepr` files included in this package.

```{r}
path <- system.file("extdata/1_1_6307.ricepr", 
                   package = "ptrapr") 

panicle <-
  path %>% 
  read_panicle()
```

We suggest that you visualize the panicle that you have loaded into R using [the package `ggraph`](https://www.ggplot2-exts.org/ggraph.html).

With a visual representation of the panicle graph, you can detect if that the XML files were parsed correctly.

For example:

```{r}
panicle %>% 
  ggraph() + 
  geom_edge_link(arrow = grid::arrow(length = unit(0.08,
                                                   "inches"),
                                     type = "closed"),
                 colour = "grey30") +
  geom_node_point(aes(colour = type),
                  size = 2,
                  alpha =.7) +
  coord_fixed()
```

## Read Grain Coordinates with `read_grains()`

Likewise, we provide a sample `.ricegr` file. This file stores coordinates for all the grains on the panicle, but no structual / topological information.

```{r}
grain_path <- system.file("extdata/1_1_6307.ricegr",
                          package = "ptrapr")

grains <- 
  grain_path %>%
  read_grains()
```

The function `read_grains()` parses and loads `.ricegr` files into R, but leaves us the challenge to integrate those points into the panicle graph. This is how they appear.

```{r}
ggplot(grains,
       aes(x = x,
           y = y)) +
  geom_point() +
  coord_fixed()
```

## Merge Grains into the Panicle Graph with `add_all_grains()`

Including grains into the panicle graph is a non trivial task with no perfect solution. We had best results by assigning the grain to the closest bracnh and by splitting the brach on the grain itself, creating a new intermediate node.

You can do this with the function `add_all_grains()`. 

```{r}
full_panicle <- 
  panicle %>%
  add_all_grains(grains)
```

For example, this is how the panicle graphs and the grains look **before** adding the grains. As you can see, the grains are detached from the main panicle graph.

```{r}
panicle %>% 
  ggraph() + 
  geom_edge_link(arrow = grid::arrow(length = unit(0.08,
                                                   "inches"),
                                     type = "closed"),
                 colour = "grey30") +
  geom_node_point(aes(colour = type),
                  size = 2,
                  alpha =.7) +
  geom_point(data = grains,
             aes(x = x, y = y)) +
  coord_fixed()
```

And this is how it looks **after adding the grains**.

```{r}
full_panicle %>%
  ggraph() + 
  geom_edge_link(arrow = grid::arrow(length = unit(0.08,
                                                   "inches"),
                                     type = "closed"),
                 colour = "grey30") +
  geom_node_point(aes(colour = type),
                  size = 2,
                  alpha =.7) +
  coord_fixed() 
```

